### 2011-09-10 paper.tex
fill order

The DP recursion is as follows (with $B,C,U,V,X,Y,u,v,x,y$ defined as in \tabref{RuleTypes})
\begin{eqnarray*}
\lefteqn{\m(i,j,k,l,A) =} \\
& &
\sum_{\rho \in \rulesubset{1}{A}} \sum_{m=i}^j W(\rho) \m(m,j-|v|,k+|x|,l-|y|,C) \mtip(i,m,U) \jvmatch \kxmatch \lymatch \\
& & +
\sum_{\rho \in \rulesubset{2}{A}} \sum_{n=k}^l W(\rho) \m(i+|u|,j-|v|,k+|x|,n,C) \mtip(n,l,Y) \iumatch \jvmatch \kxmatch \\
& & +
\sum_{\rho \in \rulesubset{3}{A}} \sum_{m=i}^j W(\rho) \m(i+|u|,m,k+|x|,l-|y|,C) \mtip(m,j,V) \iumatch \kxmatch \lymatch \\
& & +
\sum_{\rho \in \rulesubset{4}{A}} \sum_{n=k}^l W(\rho) \m(i+|u|,j-|v|,n,l-|y|,C) \mtip(k,n,X) \iumatch \jvmatch \lymatch \\
& & +
\sum_{\rho \in \rulesubset{5}{A}} \sum_{m=i}^j \sum_{n=k}^l W(\rho) \m(i+|u|,m,n,l-|y|,B) \m(m,j-|v|,k+|x|,n,C) \iumatch \jvmatch \kxmatch \lymatch \\
& & +
\sum_{\rho \in \rulesubset{6}{A}} W(\rho) \m(i+|u|,j-|v|,k+|x|,l-|y|,C)
 \iumatch \jvmatch \kxmatch \lymatch
\end{eqnarray*}

@@

The absence of null cycles implies that we can perform a topological sort of $\nodelabels$ based on the transition graph,
yielding the (reverse) order in which node labels must be visited in the Inside algorithm.

>>>>

The DP recursion relation is as follows (with $B,C,U,V,X,Y,u,v,x,y$ defined as in \tabref{RuleTypes})
\begin{eqnarray*}
\lefteqn{\m(i,j,k,l,A) =} \\
& &
\sum_{\rho \in \rulesubset{1}{A}} \sum_{m=i}^j W(\rho) \m(m,j-|v|,k+|x|,l-|y|,C) \mtip(i,m,U) \jvmatch \kxmatch \lymatch \\
& & +
\sum_{\rho \in \rulesubset{2}{A}} \sum_{n=k}^l W(\rho) \m(i+|u|,j-|v|,k+|x|,n,C) \mtip(n,l,Y) \iumatch \jvmatch \kxmatch \\
& & +
\sum_{\rho \in \rulesubset{3}{A}} \sum_{m=i}^j W(\rho) \m(i+|u|,m,k+|x|,l-|y|,C) \mtip(m,j,V) \iumatch \kxmatch \lymatch \\
& & +
\sum_{\rho \in \rulesubset{4}{A}} \sum_{n=k}^l W(\rho) \m(i+|u|,j-|v|,n,l-|y|,C) \mtip(k,n,X) \iumatch \jvmatch \lymatch \\
& & +
\sum_{\rho \in \rulesubset{5}{A}} \sum_{m=i}^j \sum_{n=k}^l W(\rho) \m(i+|u|,m,n,l-|y|,B) \m(m,j-|v|,k+|x|,n,C) \iumatch \jvmatch \kxmatch \lymatch \\
& & +
\sum_{\rho \in \rulesubset{6}{A}} W(\rho) \m(i+|u|,j-|v|,k+|x|,l-|y|,C)
 \iumatch \jvmatch \kxmatch \lymatch
\end{eqnarray*}

@@

A {\em Lexicalized TAG} (LTAG) has the property that one of $u,v,x,y$ is always a nonempty terminal string,
guaranteeing that it contains no null cycles,
and ensuring that every rule explicitly matches some output characters.
LTAGs are popular in linguistics, though the constraint is possibly too stringent to be helpful for bioinformatics grammars.

\subsubsection{Fill order}

The absence of null cycles implies that we can perform a {\em topological sort} (topo-sort) of $\nodelabels$ based on the transition graph,
yielding the order in which node labels must be visited in the Inside algorithm.

A suitable (but non-unique) Inside fill order is as follows:
\begin{itemize}
 \item For $L_1 = 0$ to $|Z|$ (increasing), \quad $i = 0$ to $|Z|-L_1$ (increasing):
 \begin{itemize}
  \item Let $l = i + L_1$
  \item For $A \in \nodelabels$ (topo-sorted, transition destinations before sources):
  \begin{itemize}
   \item For $L_2 = L_1$ to $0$ (decreasing), \quad $j = i$ to $i + L_1 - L_2$ (increasing):
   \item Let $k = j + L_2$
   \item Compute $\m(i,j,k,l,A)$ and store
  \end{itemize}
  \item Compute $\mtip(i,l,A)$ and store
 \end{itemize}
\end{itemize}
